The 2- and 3-point surface correlation functions calculations: from novel exact continuous approach to improving methodology for discrete images

### 2023-05-19 paper.tex
Addition from Alexey Kh.

\section{Introduction}
\label{sec:intro}
Correlation functions (CFs) to describe structures arose from scattering
experiments \cite{debye1957scattering} and found their way into numerous
disciplines and applications: from material sciences
\cite{Cecen,chen2022quantifying}, to rock \cite{ledesma2018effect} and soil
physics \cite{Euras2012,PLoS_ONE,KarsaninaEJSS}, cosmology \cite{TakadaJain} and
food engineering \cite{Derossi2019}. In particular, correlation functions as
computed from structural images were used for:
\begin{enumerate}
  \item characterization of morphology \cite{tensorPRE} and representativeness
    via correlation lengths \cite{vcapek2011transport,thovert2011grain} at which
    CFs reach plateau;
  \item comparison of different structures \cite{jiao2014chawla,EPL1,REVpaper},
    usually by computing L-2 norm difference between CFs for two or more images,
  \item compressing structural information \cite{SciRep1,Havelka} by
    representing huge datasets in the form of a limited set of CFs that can be
    further squeezed into parameters for a number of basis functions
    \cite{jiao2007,KarsaninaEJSS},
  \item describing structural dynamics and evolution
    \cite{chen2015dynamic,PLoS_ONE,xu2022correlation},
  \item extracting features for deep learning
    \cite{Miao2017,kamrava2020linking,roding2020predicting,KarsaninaEJSS},
  \item performing stochastic reconstructions
    \cite{Adler_recon,Y-T,EPL2,tahmasebiPRL,karsaninaPRL} based on limited input
    data such as, for example, CFs obtained from 2D cuts,
  \item fusing multi-scale images \cite{SciRep1,chen2016stochastic,Geoderma2018}
    into a single representation,
  \item evaluation of image replica quality, e.g., stochastic reconstructions
    \cite{capek2009},
  \item assessing stationarity of the structure at hand \cite{REVpaper,EfimEPL}.
\end{enumerate}

@@

@@

Now we define two disjoint sets in a homogeneous porous medium: a set of solid
phase $S$ and a set of void phase $V$. Surface-surface correlation function
$F_{ss}$ and surface-void correlation function $F_{sv}$ are then defined as
follows: \cite{Torquato_book}
\begin{align}
  F_{ss}(\bm{r}) &= \langle M_S(\bm{x}) M_S(\bm{x} + \bm{r}) \rangle \label{eq:fss} \\
  F_{sv}(\bm{r}) &= \langle M_S(\bm{x}) \chi_V(\bm{x} + \bm{r}) \rangle
  \label{eq:fsv}
\end{align}
Here the brackets $\langle \dots \rangle$ denote ensemble average. If the medium
under consideration is also isotropic, the argument in functions $F_{ss}$ and
$F_{sv}$ can be replaced with a scalar $r = |\bm{r}|$:
\begin{align}
  F_{ss}(r) &= F_{ss}(\bm{r}) \\
  F_{sv}(r) &= F_{sv}(\bm{r})
\end{align}
In case correlation functions are computed in predefined directions, scalar
values can be also used for anisotropic structures
\cite{jiao2014chawla,EPL1}. These functions are defined for media with
dimensionality of 2 and higher. $F_{sv}$ is a physical quantity inversely
proportional to length and $F_{ss}$ is inversely proportional to surface
area. SI units of measurement for these functions are $m^{-1}$ for $F_{sv}$
and $m^{-2}$ for $F_{ss}$. In practice non-standard units of measurement can be
used instead.

@@

\bibliography{pape}

>>>>

\section{Introduction}
\label{sec:intro}
Correlation functions (CFs) to describe structures arose from scattering
experiments \cite{debye1957scattering} and found their way into numerous
disciplines and applications: from material sciences
\cite{Cecen,chen2022}, to rock \cite{ledesma2018effect} and soil
physics \cite{Euras2012,PLoS_ONE,KarsaninaEJSS}, cosmology \cite{TakadaJain} and
food engineering \cite{Derossi2019}. In particular, correlation functions as
computed from structural images were used for:
\begin{enumerate}
  \item characterization of morphology \cite{tensorPRE} and representativeness
    via correlation lengths \cite{vcapek2011transport,thovert2011grain} at which
    CFs reach plateau;
  \item comparison of different structures \cite{10.1063/1.4867611,EPL1,REVpaper},
    usually by computing L-2 norm difference between CFs for two or more images,
  \item compressing structural information \cite{SciRep1,Havelka} by
    representing huge datasets in the form of a limited set of CFs that can be
    further squeezed into parameters for a number of basis functions
    \cite{jiao2007,KarsaninaEJSS},
  \item describing structural dynamics and evolution
    \cite{PhysRevE.92.023301,PLoS_ONE,xu2022correlation},
  \item extracting features for deep learning
    \cite{Miao2017,kamrava2020linking,roding2020predicting,KarsaninaEJSS},
  \item performing stochastic reconstructions
    \cite{Adler_recon,Y-T,EPL2,tahmasebiPRL,karsaninaPRL} based on limited input
    data such as, for example, CFs obtained from 2D cuts,
  \item fusing multi-scale images \cite{SciRep1,chen2016stochastic,Geoderma2018}
    into a single representation,
  \item evaluation of image replica quality, e.g., stochastic reconstructions
    \cite{capek2009},
  \item assessing stationarity of the structure at hand \cite{REVpaper,EfimEPL}.
\end{enumerate}

@@

Heterogeneous materials possess complex microstructures that play a crucial role
in determining their macroscopic properties. To gain a deeper understanding of
these materials and accurately estimate their properties, the consideration of
higher-order correlation functions becomes essential. Several n-point
correlation functions have been proposed and developed to effectively
characterize disordered materials, ranging from second-order functions like the
radial distribution function to higher-order functions involving correlations
among three or more reference points
\cite{Jiao_2012,PhysRevB.42.4453,10.1063/1.4865966}.

While certain cases may be adequately addressed by two-point correlation
functions \cite{LIU2015177,PhysRevE.77.031135}, situations demanding a higher
degree of accuracy require the higher-order correlations to construct precise
models. These higher-order correlation functions, although computationally
demanding and analytically intricate, yield valuable information and estimations
for various material properties. For instance, the effective diffusion
coefficient, a crucial transport property, can be approximated by considering a
three-point microstructural parameter expressed in terms of integrals over the
three-point probability function \cite{Jiao_2012}. Furthermore, the combination
of such functions can serve as target functions for reconstructing models of
disordered materials, enhancing the accuracy of the reconstruction process
\cite{JIAO20133370,PhysRevE.92.023301,10.1063/1.4867611,chen2016stochastic}.

Analytical approaches have been proposed to compute only n-point functions $S_n$
\cite{PhysRevB.42.4453,10.1063/1.4865966}. but studying n-point surface
correlations faces lack of analytical solutions. Availability of analytical
solutions for correlation functions holds significant importance and serves as a
desirable benchmark for evaluating and validating numerical algorithms. However,
a comprehensive computational framework for calculating higher-order correlation
functions in heterogeneous media is still lacking. In their study
\cite{malmir2018}, the authors proposed an algorithm for computing third-order
correlation functions and compared its performance with existing analytical
solutions and asymptotic cases. Although the algorithm demonstrates good
performance for the computation of S3 functions, the authors themselves
acknowledge that there is room for further improvement, particularly in the case
of surface correlation functions. To address this limitation, they suggest
employing a method similar to the one used in the work of Ma and Torquato
\cite{ma2018SS}, where a digitized image is converted into a scalar field. In
our current study, we advance beyond these findings and significantly enhance
the numerical solution for the three-point correlation function of surfaces by
incorporating and developing ideas from the work of Samarin et al
\cite{Samarin}.

@@

Now we define two disjoint sets in a homogeneous porous medium: a set of solid
phase $S$ and a set of void phase $V$. Surface-surface correlation function
$F_{ss}$ and surface-void correlation function $F_{sv}$ are then defined as
follows: \cite{Torquato_book}
\begin{align}
  F_{ss}(\bm{r}) &= \langle M_S(\bm{x}) M_S(\bm{x} + \bm{r}) \rangle \label{eq:fss} \\
  F_{sv}(\bm{r}) &= \langle M_S(\bm{x}) \chi_V(\bm{x} + \bm{r}) \rangle
  \label{eq:fsv}
\end{align}
Here the brackets $\langle \dots \rangle$ denote ensemble average. If the medium
under consideration is also isotropic, the argument in functions $F_{ss}$ and
$F_{sv}$ can be replaced with a scalar $r = |\bm{r}|$:
\begin{align}
  F_{ss}(r) &= F_{ss}(\bm{r}) \\
  F_{sv}(r) &= F_{sv}(\bm{r})
\end{align}
In case correlation functions are computed in predefined directions, scalar
values can be also used for anisotropic structures
\cite{10.1063/1.4867611,EPL1}. These functions are defined for media with
dimensionality of 2 and higher. $F_{sv}$ is a physical quantity inversely
proportional to length and $F_{ss}$ is inversely proportional to surface
area. SI units of measurement for these functions are $m^{-1}$ for $F_{sv}$
and $m^{-2}$ for $F_{ss}$. In practice non-standard units of measurement can be
used instead.

@@

\bibliography{paper}
